---
title: "scran"
author: "Zhang Yaru"
date: "2020/4/1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

## scran
* [official vignette](https://bioconductor.org/packages/release/bioc/vignettes/scran/inst/doc/scran.html)
* The scran package implements methods to perform low-level processing of scRNA-seq data, including cell cycle phase assignment, scaling normalization, variance modelling and testing for corrrelated genes. 
* Normalization by deconvolution


```{r,eval=FALSE}
suppressPackageStartupMessages(library("scran"))
suppressPackageStartupMessages(library("scater"))
```


## load data

* example data: [GSE75748](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE75748)
* date: 2016
* number of cells: 1018
* cell types: H1 H9 NPC EC DEC HFF TB
* The counts used arised from RSEM, called expected counts.(not integrate)

```{r}
data = readRDS('../../data/GSE75748/stemCell_1018.rds')
counts = data$expected_counts
counts = apply(counts,2,function(x) {storage.mode(x) = 'integer'; x})
TPM = data$TPM
```

## prepare data as SingleCellExperiment object
```{r}
sc = SingleCellExperiment::SingleCellExperiment(
  assays = list(counts = counts),
  colData = as.data.frame(data$cell_type)
)
```


## Normalization based on gene counts
* `computeSumFactors` implements the deconvolution strategy for scaling normalization
* For larger data sets, clustering should be performed with the quickCluster function before normalization.
```{r}
clusters = scran::quickCluster(sc)
sc = scran::computeSumFactors(sc, clusters=clusters)
summary(SingleCellExperiment::sizeFactors(sc))
```


* log-transform normalization using deconvolution size factors 
```{r}
sc = scater::logNormCounts(sc,size_factors=SingleCellExperiment::sizeFactors(sc))
```

* log-transfrom normalization using library size factors
```{r}
lib_sc = scater::librarySizeFactors(sc)
sc2 = scater::logNormCounts(sc,size_factors=lib_sc)
```

## relationship between library size factors and deconvolution size factors
* this two form of normalization capture similar bias.
```{r}
plot(lib_sc, SingleCellExperiment::sizeFactors(sc), xlab="Library size factor",
    ylab="Deconvolution size factor", pch=16,
    col=as.integer(factor(sc$`data$cell_type`)))
abline(a=0, b=1, col="red")
```


